---
title: "Google Capstone project"
output: html_notebook
---
Setting the language to English
```{r}
Sys.setlocale("LC_TIME", "English")
```
Installing and loading needed packages
```{r}
install.packages("tidyverse")
library("tidyverse")
install.packages("janitor")
library("janitor")
install.packages("lubridate")
library(lubridate)
```
Importing data sets
```{r}
apr_2020 <- read.csv("202004-divvy-tripdata.csv")
may_2020 <- read.csv("202005-divvy-tripdata.csv")
jun_2020 <- read.csv("202006-divvy-tripdata.csv")
jul_2020 <- read.csv("202007-divvy-tripdata.csv")
aug_2020 <- read.csv("202008-divvy-tripdata.csv")
sep_2020 <- read.csv("202009-divvy-tripdata.csv")
oct_2020 <- read.csv("202010-divvy-tripdata.csv")
nov_2020 <- read.csv("202011-divvy-tripdata.csv")
dec_2020 <- read.csv("202012-divvy-tripdata.csv")
jan_2021 <- read.csv("202101-divvy-tripdata.csv")
feb_2021 <- read.csv("202102-divvy-tripdata.csv")
mar_2021 <- read.csv("202103-divvy-tripdata.csv")
apr_2021 <- read.csv("202104-divvy-tripdata.csv")
may_2021 <- read.csv("202105-divvy-tripdata.csv")
jun_2021 <- read.csv("202106-divvy-tripdata.csv")
jul_2021 <- read.csv("202107-divvy-tripdata.csv")

```
Checking if our data sets can be put together
```{r}
compare_df_cols_same(
  apr_2020,may_2020,jun_2020,jul_2020,aug_2020,sep_2020,oct_2020,nov_2020,dec_2020,jan_2021,feb_2021,mar_2021,apr_2021,may_2021,jun_2021,jul_2021,
  bind_method = c("bind_rows", "rbind"),
  verbose = TRUE
)
```
At the current status, we can't put our data sets together as in the first 8 data sets, end_station_id and start_station_id have an inconsistent data type.
Changing them into character
```{r}

apr_2020 <- mutate(apr_2020,start_station_id = as.character(start_station_id), end_station_id = as.character(end_station_id))
may_2020 <- mutate(may_2020,start_station_id = as.character(start_station_id), end_station_id = as.character(end_station_id))
jun_2020 <- mutate(jun_2020,start_station_id = as.character(start_station_id), end_station_id = as.character(end_station_id))
jul_2020 <- mutate(jul_2020,start_station_id = as.character(start_station_id), end_station_id = as.character(end_station_id))
aug_2020 <- mutate(aug_2020,start_station_id = as.character(start_station_id), end_station_id = as.character(end_station_id))
sep_2020 <- mutate(sep_2020,start_station_id = as.character(start_station_id), end_station_id = as.character(end_station_id))
oct_2020 <- mutate(oct_2020,start_station_id = as.character(start_station_id), end_station_id = as.character(end_station_id))
nov_2020 <- mutate(nov_2020,start_station_id = as.character(start_station_id), end_station_id = as.character(end_station_id))

```
Checking if our data sets can be put together now 
```{r}
compare_df_cols_same(
  apr_2020,may_2020,jun_2020,jul_2020,aug_2020,sep_2020,oct_2020,nov_2020,dec_2020,jan_2021,feb_2021,mar_2021,apr_2021,may_2021,jun_2021,jul_2021,
  bind_method = c("bind_rows", "rbind"),
  verbose = TRUE
)
```
Putting our data sets together
```{r}
cyclistic_clean  <- rbind (apr_2020,may_2020,jun_2020,jul_2020,aug_2020,sep_2020,oct_2020,nov_2020,dec_2020,jan_2021,feb_2021,mar_2021,apr_2021,jun_2021,jul_2021)
```
Removing the columns that we are not going to use in the analysis
```{r}
cyclistic <- cyclistic_clean[ -c(9:12) ]
```
Removing rows contating NA and NULL values
```{r}
cyclistic[cyclistic == '' ] <- NA
cyclistic<- na.omit(cyclistic)
```
Calculating the length of every ride
```{r}
ride_minutes <- difftime(cyclistic$ended_at,cyclistic$started_at, units = "mins")
cyclistic <- mutate(cyclistic,ride_length = ride_minutes)
```
Adding a column with the starting hour of the ride, a column with the week of the ride and a column with the month of the ride
```{r}
hour_ride <- hour(cyclistic$started_at)
cyclistic <- mutate(cyclistic, ride_time = hour_ride)

day_ride <- wday(cyclistic$started_at, TRUE)
cyclistic <- mutate(cyclistic, ride_day = day_ride)

month_ride <- month(cyclistic$started_at)
cyclistic <- mutate(cyclistic, ride_month = month_ride)

```
If any, excluding negative or very short rides
```{r}
cyclistic <- filter(cyclistic, ride_length >= 1)

```
Exporting the data frame to .csv
```{r}
write.csv(cyclistic, "cyclistic.csv")
```